Manuscript figures

### 2025-09-04 manuscriptFigures.Rmd
added figs 8-10

@@

@@

@@

# Figure 11

>>>>

HRU low and high for 2016
HRU high = 18
HRU low = 19

@@

flow <- read_csv('./38HRU_dec2024_updates/discharge_all.csv') %>%
  replace(.==-9999, NA) %>%
  rename(date = time) %>%
  mutate(date = mdy(date)) %>%
  drop_na()

flow_pivot <- flow %>%
  pivot_longer(!date, names_to = 'cal_run', values_to = 'q_m3s') %>%
  mutate(year = as.factor(year(date))) %>%
  filter(cal_run %in% c('grid_all',
                        'statn_all',
                        'obs'
                        #'statn_2016',
                        #'grid_2016'
                        )) %>%
  mutate(cal_run = case_match(cal_run,
                            'grid_all' ~ 'Gridded',
                            'statn_all' ~ 'Station',
                            'obs' ~ 'Observed'
                          # 'statn_2016' ~ 'statn_2016',
                          #  'grid_2016' ~ 'grid_2016'
  ))

# 2016 hydrograph
flowEg <- flow_pivot %>%
  filter(year %in% c(2016)) %>%
  ggplot(aes(x = date,
             y = q_m3s, 
             color = cal_run,
             linetype = cal_run)) + 
  geom_line(size = 1, alpha = 1) + 
  theme_bw(base_size = 20) +
  scale_color_manual(values = c("Gridded" = "#0047AB", 
                                "Observed" = "black", 
                                "Station" = "orange"
                               # 'statn_2016' = 'orange',
                              #  'grid_2016' = '#0047AB'
                                )) +
  scale_linetype_manual(values = c("Gridded" = "solid",
                                   "Observed" = "dashed", 
                                   "Station" = "solid"
                                   #'statn_2016' = 'twodash',
                               #'grid_2016' = 'twodash'
)) +
  labs(x = element_blank(),
       y = "Streamflow (cms)") +
   scale_x_date(date_labels = "%b",
               date_breaks = "1 month") +
  theme(
    legend.position = c(0.9, 0.85),
    legend.text = element_text(size = 14),
    legend.key.width = unit(2, "lines"),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 2, r = 20, b = 2, l = 10)
  )

flowEg

# 2016 flow partition
gw_gridall <- read_csv(here('38HRU_dec2024_updates/catchment_output/grid_all.csv'), skip = 3) %>%
  slice(-c(1,2)) %>%
  dplyr::select(time,
                catchmentRD1_w, 
                catchmentRD2_w, 
                catchmentRG1_w, 
                catchmentRG2_w) %>%
  mutate(source = 'grid_all') %>%
  mutate(across(c(catchmentRD1_w, 
                  catchmentRD2_w, 
                  catchmentRG1_w, 
                  catchmentRG2_w), 
                as.numeric))

example <- gw %>%
  filter(year == 2016) %>%
  ungroup() %>% 
  dplyr::select(date, source,
         RD1cumsum,
         RD2cumsum,
         RG1cumsum,
         RG2cumsum) %>%
  pivot_longer(!c(date,source), names_to='flow_type',
               values_to='flow_cms') %>%
  mutate(source = if_else(source == 'grid_all',
                          'Gridded',
                          'Station'))

# some 2016 stats
flow16Stats <- example %>%
  filter(date == '2016-12-31') %>%
  group_by(source) %>%
  mutate(sum = sum(flow_cms)) %>%
  mutate(percTot = (flow_cms/sum)*100) %>%
  mutate(percTot = round(percTot, 2))

facet_labels <- c('yr_cum_RD1' = 'Surface runoff',
                  'yr_cum_RD2' = 'Interflow',
                  'yr_cum_RG1' = 'Fast groundwater',
                  'yr_cum_RG2' = 'Slow groundwater')

example_yr <- ggplot(example,
       aes(x=date, y = flow_cms, color=source)) +
  geom_line() +
  facet_wrap(~flow_type, 
             scales = 'free_y',
             labeller = labeller(flow_type = facet_labels)) +
   geom_line(size=1, alpha=0.7) +  
  scale_color_manual(values = c("#0047AB", "orange")) +
    scale_x_date(date_labels = "%b", 
               date_breaks = "2 month") +
  labs(x=element_blank(),
       y='mm of water') +
  theme_bw(base_size = 16) +
  theme(
    legend.position = 'none',
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 2, r = 20, b = 10, l = 10)
  )

example_yr

plot_grid(flowEg,
          example_yr,
          ncol=1)

ggsave('./figures/figure8.png',
       dpi=800,
       width=10,
       height=8)

@@

# params that were cal w/ burned areas
combo <- read_csv('./38HRU_dec2024_updates/params_burn_unburn.csv')

# read in pbias and nse cats
pbias <- read_csv('./38HRU_dec2024_updates/pbiasCats.csv') %>%
  filter(year == c('All')) %>%
  dplyr::select(input_data, year_missing, pbias)

# assign categories
perf <- perf %>%
  mutate(performance = case_when(
                                nse < 0.5 & abs(pbias) >= 25 ~ 'Neither',
                                nse > 0.5 & abs(pbias) >= 25 ~ 'NSE',
                                nse < 0.5 & abs(pbias) < 25 ~ '% bias satisfactory',
                                nse > 0.5 & abs(pbias) < 25 ~ 'NSE + % bias satisfactory')) %>%
  mutate(input_data = if_else(input_data == 'grid',
                              'Gridded', 'Station'),
         year_missing = if_else(year_missing == 'None',
                                'all', year_missing))
perf1 <- perf %>%
  mutate(year_missing = if_else(year_missing == 'all',
                                'none',
                                year_missing))

combo <- left_join(combo,
                        perf1,
                        by = c('input_data',
                               'year_missing'))

# reorder the params
# combo$param <- factor(combo$param, levels = c("a_snow",
#                                                         "a_rain",
#                                                         "soilMaxInfSummer",
#                                                         "soilMaxDPS",
#                                                         "lagSurfaceRunoff",
#                                                         "soilDistMPSLPS"))
combo <- combo %>%
  mutate(param = fct_recode(param,
    "a_rain (mm)" = "a_rain",
    "a_snow (mm)" = "a_snow",
    "soilMaxInfSummer (mm/day)" = "soilMaxInfSummer",
    "soilMaxDPS (mm)" = "soilMaxDPS",
    "lagSurfaceRunoff" = "lagSurfaceRunoff",
    "soilDistMPSLPS" = "soilDistMPSLPS"
  ))

# text above boxplots - this is from Tukey analysis that was already run below then added to params
combo <- combo %>%
  group_by(param, burn, input_data) %>%
  mutate(y_text = max(value, na.rm = TRUE) * 1.08) %>%  # Slightly above max value
  ungroup()

ggplot(combo, aes(x = burn, y = value, fill = input_data)) +
  geom_boxplot(alpha = 0.8, 
               outliers = F) +
  geom_text(aes(label = letter, y = y_text), 
            position = position_dodge(width = 0.75),  # Align with boxplots
            size = 6, 
            vjust = 0) +  
  geom_jitter(aes(shape = performance),
            position = position_jitterdodge(
    jitter.width = 0.1,
    dodge.width = 0.75),
            size = 2.5, 
            alpha = 0.6,
            stroke = 1,
            color = "black") +
  scale_shape_manual(values = c(
  "% bias satisfactory" = 1,
  "NSE + % bias satisfactory" = 16)) +
  facet_wrap(~param, 
             scales = 'free_y',
             strip.position = 'left') +
  labs(y = element_blank(), x = element_blank()) +
  theme_bw(base_size = 20) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
  theme(
    legend.position = 'bottom',
    legend.text = element_text(size = 16),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    strip.placement = 'outside',
    axis.title.y = element_blank()
  ) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))

ggsave('./figures/figure9.png',
       dpi=800,
       width = 12,
       height = 8)

@@

unburn <- read_csv('./38HRU_dec2024_updates/HRU/statn_HRU_2_unburn.csv') %>%
  mutate(date = mdy(time),
         year = year(date),
         type = 'unburn') %>%
  filter(year == 2016)

burn <- read_csv('./38HRU_dec2024_updates/HRU/statn_HRU_5_burn.csv') %>%
  mutate(date = mdy(time),
         year = year(date),
         type = 'burn') %>%
  filter(year == 2016)

both <- bind_rows(unburn, burn)

# interception
inter <- both %>%
  dplyr::select(date, type, intercStorage)

# swe
swe <- both %>%
  dplyr::select(date, type, snowTotSWE)

p <- both %>%
  dplyr::select(date, type, precip) %>%
  group_by(type) %>%
  mutate(p_cumu = cumsum(precip))

# soil sat
soilsat <- both %>%
  dplyr::select(date, type, soilSat)

# surface runoff
runoff <- both %>%
  dplyr::select(date, type, outRD1)

# interflow
interflow <- both %>%
  dplyr::select(date, type, outRD2) %>%
  mutate(interflow_cum = cumsum(outRD2))

gw <- both %>%
  dplyr::select(date, type, outRG1) %>%
  mutate(gw_cumu = cumsum(outRG1))

# dps
dps <- both %>%
  dplyr::select(date, type, actDPS)

# mps
mps <- both %>%
  dplyr::select(date, type, soilSatMPS)

# lps
lps <- both %>%
  dplyr::select(date, type, soilSatLPS)

# soil sat
soilsat <- both %>%
  dplyr::select(date, type, soilSat)

## make the actual plot
p_plot <- ggplot(p, aes(x = date,
                        y = p_cumu, 
                        color = type)) +
  geom_line(size=1, alpha=0.8) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Precipitation (mm)') +
  theme_bw(base_size = 18) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

p_plot

inter <- inter %>%
  mutate(type = factor(type, levels = c("unburn", "burn")))

int_plot <- ggplot(inter, aes(x = date,
                        y = intercStorage, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Interception (mm)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

int_plot

soilsat_plot <- ggplot(soilsat, aes(x = date,
                        y = soilSat, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Soil saturation') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

soilsat_plot

interflow_plot <- ggplot(interflow, aes(x = date,
                        y = interflow_cum, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Interflow (L)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

interflow_plot

gw_plot <- ggplot(gw, aes(x = date,
                        y = gw_cumu, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Shallow groundwater (L)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

gw_plot

swe_plot <- ggplot(swe, aes(x = date,
                        y = snowTotSWE, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'SWE (mm)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

swe_plot

legend <- cowplot::get_plot_component(p_plot, "guide-box", return_all = TRUE)[[3]]

p_plot <- p_plot + theme(legend.position = "none")
int_plot <- int_plot + theme(legend.position = "none")
soilsat_plot <- soilsat_plot + theme(legend.position = "none")
interflow_plot <- interflow_plot + theme(legend.position = "none")
gw_plot <- gw_plot + theme(legend.position = "none")
swe_plot <- swe_plot + theme(legend.position = "none")

plots <- plot_grid(p_plot,
          int_plot,
          swe_plot,
          soilsat_plot,
          interflow_plot,
          gw_plot,
          ncol = 3)

final_plot <- plot_grid(
  plots,
  legend,
  ncol = 1,
  rel_heights = c(1, 0.07)  # Adjust height of legend
)

ggsave('./figures/figure10.png',
       dpi=800,
       bg='white',
       height=8,
       width=12)
